\documentclass[twoside,openright]{report}

\usepackage{cclicenses}
\usepackage{emptypage}
\usepackage{fancyhdr}
\usepackage{textcomp}
\usepackage{hyperref}
\pagestyle{fancy}

\fancyfoot[C]{This document is provided for informational use only. It is for
use by authorized instructors. Consult the aircraft manual and appropriate FAA
handbooks to double-check all information. \textcopyright{} Derek W Beck 2008.
Some Rights Reserved. \url{http://www.derekbeck.com/}.
\\
\byncsa{} \href{https://creativecommons.org/licenses/by-nc-sa/3.0/us/}{Licensed
under Creative Commons Attribution-Noncommercial-Share Alike License}.
}

\fancypagestyle{plain}

\title{Certificated Flight Instructor – Airplane (CFI-A)
Private and Commercial Pilot Lesson Plans}
\author{Derek W Beck}
\date{June 2008}

\begin{document}

\makeatletter
\begin{titlepage}
  \begin{center}
    \vspace*{1cm}
    \textbf{\@title}

    \vspace*{1cm}
    \@author

    \@date

    \vspace*{2cm}
    \ccby You may copy, distribute, display this copyrighted work — and
    derivative works based upon it — but only if they give credit to Derek W
    Beck and abide by the other license requirements listed here.

    \ccnc This work and its derivatives may not be sold without permission from
    Derek W Beck.

    \ccsa You may distribute derivative works only under a license identical to
    the licenses listed here and only if these licenses are explicitly depicted
    somewhere on the derivative work.
  \end{center}
\end{titlepage}

\tableofcontents

\chapter{Hello}

This is some text.

\chapter{World}

This is some more text.

\end{document}
