\documentclass[twoside,openright]{report}

\let\oldsection\section
\def\section{\cleardoublepage\oldsection}

\usepackage[margin=1in]{geometry}
\usepackage{cclicenses}
\usepackage{emptypage}
\usepackage{fancyhdr}
\usepackage{textcomp}
\usepackage{hyperref}
\pagestyle{fancy}

\fancyfoot[C]{This document is provided for informational use only. It is for
use by authorized instructors. Consult the aircraft manual and appropriate FAA
handbooks to double-check all information. \textcopyright{} Derek W Beck 2008.
Some Rights Reserved. \url{http://www.derekbeck.com/}.
\\
\byncsa{} \href{https://creativecommons.org/licenses/by-nc-sa/3.0/us/}{Licensed
under Creative Commons Attribution-Noncommercial-Share Alike License}.
}

\fancypagestyle{plain}

\title{Certificated Flight Instructor – Airplane (CFI-A)
Private and Commercial Pilot Lesson Plans (C172RG)}
\author{Derek W Beck}
\date{June 2008}

\begin{document}

\makeatletter
\begin{titlepage}
  \begin{center}
    \vspace*{1cm}
    \textbf{\@title}

    \vspace*{1cm}
    \@author

    \@date

    \vspace*{2cm}
    \ccby You may copy, distribute, display this copyrighted work -- and
    derivative works based upon it -- but only if they give credit to Derek W
    Beck and abide by the other license requirements listed here.

    \ccnc This work and its derivatives may not be sold without permission from
    Derek W Beck.

    \ccsa You may distribute derivative works only under a license identical to
    the licenses listed here and only if these licenses are explicitly depicted
    somewhere on the derivative work.
  \end{center}
\end{titlepage}

\setcounter{tocdepth}{1}
\tableofcontents

\chapter{Ground Lessons}

\input{../common/pp-intro}
\input{../common/considerations}
\input{../common/preflight}
\input{../common/aerodynamics}
\input{../common/systems}
\input{fuel-system-c172rg}
\input{electrical-system-c172rg}
\input{landing-gear-c172rg}
\input{../common/propeller}
\input{../common/vfr-instruments}
\input{../common/docs-and-maintenance}
\input{../common/weight-and-balance}
\input{../common/weather-theory}
\input{../common/weather-services}
\input{../common/airspace}
\input{../common/airport-operations}
\input{../common/emergency-procedures}
\input{../common/visual-illusions}
\input{../common/spin-avoidance}
\input{../common/vfr-flight-planning}

\chapter{Maneuvers}

\input{accelerated-stalls-c172rg.tex}
\input{basic-instrument-maneuvers.tex}
\input{chandelles.tex}
\input{climbs.tex}
\input{descents.tex}
\input{cross-control-stalls.tex}
\input{crosswind-landings.tex}
\input{crosswind-takeoffs.tex}
\input{eights-on-pylons.tex}
\input{elevator-trim-stalls.tex}
\input{emergency-approaches-and-landings.tex}
\input{forward-slip-to-land.tex}
\input{go-arounds.tex}
\input{lazy-eights.tex}
\input{level-turns.tex}
\input{normal-landings.tex}
\input{normal-takeoffs.tex}
\input{power-off-180.tex}
\input{power-off-stalls.tex}
\input{power-on-stalls.tex}
\input{rectangular-course.tex}
\input{secondary-stalls.tex}
\input{short-field-landing.tex}
\input{short-field-takeoff.tex}
\input{slow-flight.tex}
\input{soft-field-landing.tex}
\input{soft-field-takeoff.tex}
\input{spins.tex}
\input{steep-spiral.tex}
\input{steep-turns.tex}
\input{straight-and-level.tex}
\input{s-turns.tex}
\input{turns-around-a-point.tex}
\input{unusual-attitudes.tex}

\end{document}
